Manuscript Figures for Genome Biology

### 2018-07-06 manuscript_figures/manuscript_figures.Rmd
Incorporate uninformative sb objects; standardize heatmap styles

```{r Figure2}
objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("yeast", objects) & 
                   grepl("de5", objects) & 
                  !grepl("uninfCov", objects)]

@@

# remove microbiome objects from 'null' comparisons
# removing baxter, goodrich and papa OTU-level (since they mostly found 
# nothing for all methods and the genus-level analysis is also present)
null <- which(grepl(c("goodrich|papa"), objects) & 
              grepl(c("otu"), objects) | 
              grepl("baxter", objects))

@@

@@

Fig4.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                            tableOnly = TRUE)$method)
Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                        annotate = "propPossible")
Fig4

# not excluding any methods
Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha, 
                               tableOnly = TRUE, excludeMethods = NULL)$method)
Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                        excludeMethods = NULL)
Fig4All
```

@@

# non-null
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|null|polyester"), dataset) &
            !grepl("scRNAseq", casestudy) &
            !grepl("uninfCov", dataset))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

pfdr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = alpha, xlab = "Comparison",
                fill = "FDR",
                rowOrder = Fig4.rowOrder, 
                annotate = "FDR") +
          scale_fill_gradient2(high = "red", 
                         mid = "white", 
                         low = "blue", 
                         midpoint = alpha)

ptpr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = alpha, xlab = "Comparison",
                fill = "TPR",
                colorLow = "blue", 
                colorHigh = "white",
                rowOrder = Fig4.rowOrder,
                annotate = "TPR")

ggdraw() +
  draw_plot(pfdr + ylab("") + xlab(""), 0.57, 0, 0.22, 1) +
  draw_plot(ptpr + ylab("") + xlab(""), 0.78, 0, 0.22, 1) +
  draw_plot(Fig4, 0, 0, 0.60, 1) +
  draw_plot_label(c("A", "B"), c(0, 0.57), c(1, 1), size = 15)
ggsave(file.path(outdir, "Figure4.pdf"), width=12, height=5)

@@

@@

objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pA <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
                        nbins=25) +
  ggtitle("RNA-seq case study: Brain") +
  xlab("Overall mean expression percentile")

@@

resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
                   grepl("mast", objects) & grepl("det", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pB <- covariateLinePlot(res, alpha=alpha, covname="detection", trans="log1p",
                        nbins=25) +
  ggtitle("scRNA-seq case study: Human, MAST") +
  xlab("Detection rate percentile") + 
  theme(legend.position="none")

resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pC <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
                        nbins=25) +
  ggtitle("GSEA case study: Mouse") +
  xlab("Gene Set size percentile") + 
  theme(legend.position="none")

@@

p5 <- plot_grid(pA, pB, pC, pD, ncol=2, labels = LETTERS[1:4])
p5 <- plot_grid(p5, leg, ncol=2, rel_widths = c(1,0.2))
ggsave(file.path(outdir, "Figure5.pdf"), width=9, height=7)

@@

```{r}
# by covariate
wcs <- which(grepl("GWAS", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
                      function(x) x[2]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Covariate",
                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("GWAS")
ggsave(file.path(outdir, "FigureS1.pdf"), width=4, height=4)
```

@@

```{r}
#by dataset
wcs <- which(grepl("GSEA", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("GSEA")
ggsave(file.path(outdir, "FigureS2.pdf"), width=4, height=4)
```

@@

```{r}
wcs <- which(grepl("ChIPseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("ChIP-Seq")
ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=4)
```

### RNA-seq
```{r}
# by dataset
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("polyester|yeast"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("RNA-seq")
ggsave(file.path(outdir, "FigureS4.pdf"), width=4, height=4)
```

@@

# by dataset
pa <- plotMethodRanks(objs, colLabels = ds, alpha = alpha, 
                      xlab = "Dataset", annotate = "propPossible") + 
  ggtitle("scRNA-seq")

# by method
pb <- plotMethodRanks(objs, colLabels = meth, alpha = alpha,
                      xlab = "Method", annotate = "propPossible") + 
  ggtitle("scRNA-seq")

# by covariate
pc <- plotMethodRanks(objs, colLabels = cv, alpha = alpha,
                      xlab = "Covariate", annotate = "propPossible") + 
  ggtitle("scRNA-seq")

plot_grid(pa, pb, pc)
ggsave(file.path(outdir, "FigureS5.pdf"), width=8, height=5.5)
```

@@

```{r}
# by dataset, excluding otu when genus is available
# excluding log-ubiquity
# excluding enigma-so4 (no discoveries)
wcs <- which(grepl("microbiome", casestudy) & 
              !(grepl(c("goodrich|papa"), dataset) & 
              grepl(c("otu"), dataset)) &
              !grepl("log-ubiquity", dataset) &
              !grepl("so4", dataset) &
              !grepl("baxter", dataset) )
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("Microbiome")
ggsave(file.path(outdir, "FigureS6.pdf"), width=7.5, height=4)
```

@@

```{r}
# non-null
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|polyester|null|uninfCov"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

# null
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|polyester|de|uninfCov"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

@@

```{r}
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|yeast|null|uninfCov"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

@@

## exclude yeast and polyester simulations
objects <- objects[!grepl("yeast", objects) & !grepl("polyester", objects)]

@@

## read in case study results (proportion of rejects vs. case study max)
objdat <- tidy_df(objects, object_labs, "propMaxRejections", NULL)
objdat <- as_tibble(objdat)
objdat <- tidyr::complete(objdat, method, casestudy)

>>>>

```{r Figure2}
objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("yeast", objects) & 
                   grepl("de5", objects) & 
                  !grepl("uninf", objects)]

@@

# remove microbiome objects from 'null' comparisons
# removing baxter, goodrich, papa OTU-level, and So4-enigma (since they mostly found 
# nothing for all methods and the genus-level analysis is also present)
# Also remove'log' ubiquity since redundant
null <- which(grepl(c("goodrich|papa"), objects) & 
              grepl(c("otu"), objects) | 
              grepl("baxter", objects) |
              grepl("log", objects) | 
              grepl("so4", objects))

@@

# remove random covariate analyses
uninf <- which(grepl("uninf", objects))
if (length(uninf) > 0){
  objects <- objects[-uninf]
}

@@

Fig4.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
                   "ashq")
Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                        annotate = "propPossible")
Fig4

# not excluding any methods
Fig4All.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
                  "scott-empirical", "scott-theoretical", "ashq")
Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                        excludeMethods = NULL, rowOrder = Fig4All.rowOrder)
Fig4All
```

@@

# non-null
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|null|polyester"), dataset) &
            !grepl("scRNAseq", casestudy) &
            !grepl("uninf", dataset))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

pfdr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = alpha, xlab = "Comparison",
                fill = "FDR",
                rowOrder = Fig4All.rowOrder, 
                annotate = "FDR", excludeMethods = NULL) +
          scale_fill_gradient2(high = "red", 
                         mid = "white", 
                         low = "blue", 
                         midpoint = alpha) +
  labs(fill = "Mean FDR")

ptpr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = alpha, xlab = "Comparison",
                fill = "TPR",
                rowOrder = Fig4All.rowOrder,
                annotate = "TPR", excludeMethods = NULL)

ggdraw() +
  draw_plot(pfdr + ylab("") + xlab(""), 0.57, 0, 0.22, 1) +
  draw_plot(ptpr + ylab("") + xlab(""), 0.78, 0, 0.22, 1) +
  draw_plot(Fig4All, 0, 0, 0.60, 1) +
  draw_plot_label(c("A", "B"), c(0, 0.57), c(1, 1), size = 15)
ggsave(file.path(outdir, "Figure4.pdf"), width=14, height=5)

@@

ggplot(rank.tab , 
       aes(y = prop_casestudy, x = method, group = method, fill = Method)) +
  geom_boxplot(alpha=0.7) +
  scale_fill_manual(values=col) +
  theme_bw() +
  xlab("Method (ordered by rejections in yeast simulation (n=5))") +  
  ylab("Proportion of maximum rejections in casestudies")
ggsave(file.path(outdir, "Figure_simVcasestudyRanks.pdf"), width=6, height=4)

@@

objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects) & !grepl("uninf", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pA <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
                        nbins=25) +
  ggtitle("RNA-seq case study: Brain") +
  xlab("Overall mean expression percentile")

@@

resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
                   grepl("mast", objects) & grepl("det", objects) & !grepl("uninf", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pB <- covariateLinePlot(res, alpha=alpha, covname="detection", trans="log1p",
                        nbins=25) +
  ggtitle("scRNA-seq case study: Human, MAST") +
  xlab("Detection rate percentile") + 
  theme(legend.position="none")

resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pC <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
                        nbins=25) +
  ggtitle("GSEA case study: Mouse") +
  xlab("Gene Set size percentile") + 
  theme(legend.position="none")

@@

p5 <- plot_grid(pA, pB, pC, pD, ncol=2, labels = LETTERS[1:4])
p5 <- plot_grid(p5, leg, ncol=2, rel_widths = c(1,0.2))
p5
ggsave(file.path(outdir, "Figure5.pdf"), width=9, height=7)

@@

```{r}
# by covariate
wcs <- which(grepl("GWAS", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
                      function(x) x[2]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Covariate",
                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("GWAS") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
ggsave(file.path(outdir, "FigureS1.pdf"), width=5, height=4.5)
```

@@

```{r}
#by dataset
wcs <- which(grepl("GSEA", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("GSEA") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
ggsave(file.path(outdir, "FigureS2.pdf"), width=6, height=4.5)
```

@@

```{r}
wcs <- which(grepl("ChIPseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("ChIP-Seq") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=5)
```

### RNA-seq
```{r}
# by dataset
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("polyester|yeast"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("RNA-seq") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
ggsave(file.path(outdir, "FigureS4.pdf"), width=6, height=5)
```

@@

uninf <- which(grepl("uninf", objs))

# by dataset
pa <- plotMethodRanks(objs[-uninf], colLabels = ds[-uninf], alpha = alpha, 
                      xlab = "Dataset", annotate = "propPossible") + 
  ggtitle("scRNA-seq")+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))

# by method
pb <- plotMethodRanks(objs[-uninf], colLabels = meth[-uninf], alpha = alpha,
                      xlab = "Method", annotate = "propPossible") + 
  ggtitle("scRNA-seq")+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))

# by covariate
pc <- plotMethodRanks(objs, colLabels = cv, alpha = alpha,
                      xlab = "Covariate", annotate = "propPossible") + 
  ggtitle("scRNA-seq")+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))

plot_grid(pa, pb, pc, ncol=3)
ggsave(file.path(outdir, "FigureS5.pdf"), width=15, height=6)
```

@@

```{r}
# by dataset, excluding otu when genus is available
# excluding log-ubiquity
# excluding enigma-so4 (no discoveries)
wcs <- which(grepl("microbiome", casestudy) & 
              !(grepl(c("goodrich|papa"), dataset) & 
              grepl(c("otu"), dataset)) &
              !grepl("log-ubiquity", dataset) &
              !grepl("so4", dataset) &
              !grepl("baxter", dataset) )
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                Nlabel = FALSE, annotate = "propPossible") + 
  ggtitle("Microbiome")+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
ggsave(file.path(outdir, "FigureS6.pdf"), width=14, height=6.5)
```

@@

```{r}
# non-null
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|polyester|null"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL,
                annotate = "propPossible")

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL, 
                fill = "FDR",
                annotate = "propPossible")

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL, 
                fill = "TPR",
                annotate = "propPossible")

# null
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|polyester|de"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL,
                annotate = "propPossible")

@@

```{r}
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("brain|mir|yeast|null"), objects) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- gsub(".rds", "", 
             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
                      function(x) x[2])))

@@

## exclude yeast and polyester simulations, as well as uninformative covariate
## analyses
objects <- objects[!grepl("yeast", objects) & !grepl("polyester", objects) &
                   !grepl("uninf", objects)]

@@

## read in case study results (proportion of rejects vs. case study max)
objdat <- tidy_df(objects=objects, colLabels=object_labs, 
                  fill="propMaxRejections", annotate=NULL,
                  alpha =0.10)
objdat <- as_tibble(objdat)
objdat <- tidyr::complete(objdat, method, casestudy)
